Moral Foundations of U.S. Political News Organizations

### 2021-12-13 thesis_version/pilot_manuscript.Rmd
cleaning up

abstract: |
  The media ecosystem has grown, and political opinions have diverged such that there are competing conceptions of objective truth. Commentators often point to political biases in news coverage as a catalyst for this political divide. The Moral Foundations Dictionary (MFD) facilitates identification of ideological leanings in text through frequency of the occurrence of certain words. Through web scraping, the researchers extracted articles from popular news sources' websites, calculated MFD word frequencies, and identified words' respective valences. This process attempts to uncover news outlets' positive or negative endorsements of certain moral dimensions concomitant with a particular ideology. In Experiment 1, the researchers gathered poltical articles from four sources. We were unable to reveal significant differences in moral or political endorsements, but we solidified the method to be employed in further research. In Experiment 2, the researchers expanded their number of sources to 10 and analyzed articles that pertain to two specific topics: the 2018 confirmation hearings of U.S. Supreme Court Justice Brett Kavanaugh and the partial U.S. Goverment Shutdown of 2018-2019. Once again, no significant differences in moral or political endorsements were found.

@@

Borrowing from Osgood's work in the 1950s, @Bradley1999 recognized valence as one of three related dimensions comprising emotion when developing their Affective Norms for English Words (ANEW). As mentioned before, "valence," the first dimension, denotes the pleasantness of a given word. "Arousal," the second dimension, describes the stimulating nature of a word. Lastly, "dominance" or "control" describes the extent to which a word makes one feel in or out of control [@Bradley1999]. The researchers developed ANEW by presenting participants with a list of 100-150 words and asking for them to rate the word on all three dimensions using the Self-Assessment Mannikin (SAM), which allows ratings along either a nine-point scale when using traditional paper instruments or a twenty-point scale when using a computerized version.

@@

##import the overalldata 
overalldata = read.csv("exp1_data/overalldata.csv")

@@

```{r data-cleanup, eval=FALSE, include=FALSE}
##pull in the data
master = read.csv("exp1_data/overalldata.csv", stringsAsFactors = F)

@@

final = read.csv("exp1_data/finaldata.csv", stringsAsFactors = F)
final = na.omit(final) #one weird NA line

@@

```{r stem_warriner, include = FALSE}
##open the Warriner data for later
warriner = read.csv("exp1_data/BRM-emot-submit.csv", stringsAsFactors = F) #want v.mean.sum
warriner$V.Mean.Sum = scale(warriner$V.Mean.Sum)

@@

```{r create_datasets_harm, echo = F, include=FALSE}
#read in the harm total words from the mtmm project
harm_all = read.csv("exp1_data/harm_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_fair, include=FALSE}
fair_all = read.csv("exp1_data/fair_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_ingroup, include=FALSE}
ingroup_all = read.csv("exp1_data/ingroup_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_authority, include=FALSE}
authority_all = read.csv("exp1_data/authority_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_purity, include=FALSE}
purity_all = read.csv("exp1_data/purity_words_total.csv", stringsAsFactors = F)

@@

As seen in Table \@ref(tab:exp1-source-descriptives), the sources are similar in some aspects yet different in others. Valence appears to be slightly positive across all sources. The large standard deviations seem to indicate little to no presence of a difference in valence across sources. 
*The New York Times* published the greatest number of articles as well as total words. *Breitbart* featured the lowest number of articles, and *NPR* the lowest number of total words from all articles. Per individual article, however, *Breitbart* appears to feature the highest average number of words as well as unique words. Once again the standard deviations call into question any apparent differences between sources. Finally, *Fox News* articles had the lowest reading grade level on average, while *The New York Times* ihad the highest. This result might be attributable to the greater number of tokens in the average *New York Times* article compared to *Fox News*. The standard deviations for readability suggest the presence of a diverse array of articles for each source, ranging from low to high reading level. Large standard deviations suggest the sources feature a lot of overlap between them in their representation of scores.

@@

```{r data_clean_kav, eval = T, echo = F}
##combine all kav data into one dataset
#read it in
NYtimes_kav = read.csv("exp2_data/NYtimes_kav.csv", stringsAsFactors = F, fileEncoding="UTF-8") #ok
NPR_kav = read.csv("exp2_data/NPR_kav.csv", stringsAsFactors = F, fileEncoding="UTF-8") #ok
SLATE_kav = read.csv("exp2_data/SLATE_kav.csv", stringsAsFactors = F, fileEncoding="UTF-8") #ok
HUFFPO_kav = read.csv("exp2_data/HUFFPO_kav.csv", stringsAsFactors = F, fileEncoding="UTF-8") #ok
POLITICO_kav = read.csv("exp2_data/POLITICO_kav.csv", stringsAsFactors = F, fileEncoding="UTF-8") #ok 
FOX_kav = read.csv("exp2_data/FOX_kav.csv", stringsAsFactors = F) #ok 
BREITBART_kav = read.csv("exp2_data/BREITBART_kav.csv", stringsAsFactors = F) #ok
RUSH_kav = read.csv("exp2_data/RUSH_kav.csv", stringsAsFactors = F, fileEncoding="UTF-8") #ok 
BLAZE_kav = read.csv("exp2_data/BLAZE_kav.csv", stringsAsFactors = F, fileEncoding="UTF-8") #ok 
HANNITY_kav = read.csv("exp2_data/HANNITY_kav.csv", stringsAsFactors = F, fileEncoding="UTF-8") #ok

@@

```{r data_clean_gs, eval = T, echo = F}
##combine all gs data into one dataset
#read it in
NYtimes_gs = read.csv("exp2_data/NYtimes_gs.csv", stringsAsFactors = F, fileEncoding = "UTF-8") #ok
NPR_gs = read.csv("exp2_data/NPR_gs.csv", stringsAsFactors = F) #ok
SLATE_gs = read.csv("exp2_data/SLATE_gs.csv", stringsAsFactors = F) #ok 
HUFFPO_gs = read.csv("exp2_data/HUFFPO_gs.csv", stringsAsFactors = F) #ok 
POLITICO_gs = read.csv("exp2_data/POLITICO_gs.csv", stringsAsFactors = F, fileEncoding = "UTF-8") #not ok 
FOX_gs = read.csv("exp2_data/FOX_gs.csv", stringsAsFactors = F) #ok 
BREITBART_gs = read.csv("exp2_data/BREITBART_gs.csv", stringsAsFactors = F) #ok 
RUSH_gs = read.csv("exp2_data/RUSH_gs.csv", stringsAsFactors = F, fileEncoding = "UTF-8") #not ok 
BLAZE_gs = read.csv("exp2_data/BLAZE_gs.csv", stringsAsFactors = F) #ok 
HANNITY_gs = read.csv("exp2_data/HANNITY_gs.csv", stringsAsFactors = F) #ok

@@

```{r create_datasets_harm_kav, eval = T, echo = F, include = F}
#read in the harm total words from the mtmm project
harm_all_kav = read.csv("exp1_data/harm_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_fair_kav, include=FALSE}
fair_all_kav = read.csv("exp1_data/fair_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_ingroup_kav, include=FALSE}
ingroup_all_kav = read.csv("exp1_data/ingroup_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_authority_kav, include=FALSE}
authority_all_kav = read.csv("exp1_data/authority_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_purity_kav, include=FALSE}
purity_all_kav = read.csv("exp1_data/purity_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_harm_gs, eval = T, echo = F, include = F}
#read in the harm total words from the mtmm project
harm_all_gs = read.csv("exp1_data/harm_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_fair_gs, echo = F, include=FALSE}
fair_all_gs = read.csv("exp1_data/fair_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_ingroup_gs, echo = F, include=FALSE}
ingroup_all_gs = read.csv("exp1_data/ingroup_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_authority_gs, echo = F, include=FALSE}
authority_all_gs = read.csv("exp1_data/authority_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_purity_gs, echo = F, include=FALSE}
purity_all_gs = read.csv("exp1_data/purity_words_total.csv", stringsAsFactors = F)

@@

The results obtained in both experiments, while not confirmatory, provide several avenues for future investigation. Twice confronted with the failure to uncover differences in news media's political content, the researchers were unable to confirm the existence of an effect of political lean on American news writing under the auspices of Moral Foundations Theory. This does not, however, rule out the possibility that news media is biased. As @Mitchell2014 has demonstrated, self-reported political partisans are drawn to separate sources, but the present research was unable to offer any content-based explanations of this phenomenon. Even when attenuating the scraping criteria to collect text concerning (what the researchers believed to be) morally resonant topics, the method failed to detect differences in MFD word use by prominent news sources. Given part of the researchers' goal with this study was to answer a method question, they must contemplate whether or not the method or the construct of interest (political lean) are the primary contributor to null results. Either the method used to detect lean was insufficent for this purpose, or political lean simply does not exist, indicating the presence of a true negative in the results.

The researchers believe these results represent a false negative rather than a true negative. That is, we believe political lean exists, but that we have been unable to detect it at a content level. The researchers again turn to @Mitchell2014's work in this area: political partisans prefer certain news outlets over others. There must be something about the sources' content that facilitates partisans' repeated usage. Otherwise, they might not express such loyalty to any one source (or group of sources) over others. Given this aspect of the news economy, the researchers still believe certain news outlets in the U.S. contatin a political slant. The researchers' focus, therefore, turns toward the limitations of the method, especially the chosen theoretical framework, under which the present research was conducted.

@@

There remains an aspect of language which this method still cannot address: context. Owing to the presence of millions of words in the analyses presented in the current study, the researchers would be unable to check for context and correct endorsement scores accordingly. @Graham2009 actually performed a process of human error-checking for context when formulating the MFD. While this might have helped that team fix mistakes made by the LIWC program (@Pennebaker2007), manual coding of this nature has the potential to introduce an unnacceptable level of bias in the method, as perceptions of context would be dependent upon raters' interpretations and beleif systems. Additionally, the numerous ways in which context affects meaning make it difficult to interpret MFD word occurrence outside of a narrow operational definition of "endorsement." In fact, political lean may be communicated through methods other than through the use of particular words.

One method through which news organizations might appeal to political partisans is through their content strategy rather than their use of MFD words. The results in the current study demonstrate a similar pattern of endorsement across both conservative and liberal sources. Specifically, the researchers found slightly positive endorsements for all five foundations no matter the political lean of the source. While this may simply be an indication of insufficient MFD word occurrence in news articles, it may also demonstrate truly equal endorsement despite political lean. While news organizations may endorse all moral foundations, they may also adopt strategies for doing so that appeal to the political partisans consuming their product. As an example, a liberal source might endorse *purity/sanctity* through an appeal to reduce waste or pollution (i.e. keeping the earth "clean"). A conservative source, on the other hand, might endorse *purity/sanctity* through an argument in favor of abstinence or religious commitment (i.e. keeping a "clean" soul, mind, or body). Both publications would register as positive endorsement of *purity/sanctity* despite appropriating the concept for distinct purposes. Through these divergent strategies, liberal and conservative news organizations might be able to appeal to their fans while exhibiting similar levels of moral endorsement. This could represent a further weakness of the Moral Foudnations Dictionary as an instrument, as it is unable to identify these potential strategies for communicating certain points of view.

@@

@Belisle2018 conducted a text analysis (using a computer algorithm) of several speeches made by four recent U.S. presidents under Relational Frame Theory. These speeches, representing prominent examples of political discourse, were found to differ in their use of relational frames. For example, speeches by Donald Trump were found to contain more distinctive frames than other presidents, while Barack Obama utilized more coordinative frames than his counterparts. According to @Belisle2018, these observations imply President Obama, who empolyed more coordinative frames, utilized language emphasizing unity and similarity between people. Donald Trump, on the other hand emphasized differences and divisions between people through his use of distinctive frames [@Belisle2018]. At first glance, the differences appear to fall along party lines, though other factors may be involved as well. In any case, this research appears to indicate that RFT might be useful for analyzing political news content similar to the method employed in the current study.

@@

```{r appendix_mfd, results = 'asis', echo = FALSE, message = F}
orig_mfd = read.csv("exp1_data/original_mfd.csv", stringsAsFactors = F, fileEncoding="UTF-8")
####build the table here####
tableprint = matrix(NA, nrow=length(orig_mfd), ncol=5)
colnames(tableprint) = c("Harm/Care", "Fairness/Reciprocity", "Ingroup/Loyalty", "Authority/Respect", "Purity/Sanctity")

>>>>

abstract: |
  The media ecosystem has grown, and political opinions have diverged such that there are competing conceptions of objective truth. Commentators often point to political biases in news coverage as a catalyst for this political divide. The Moral Foundations Dictionary (MFD) facilitates identification of ideological leanings in text through frequency of the occurrence of certain words. Through web scraping, the researchers extracted articles from popular news sources' websites, calculated MFD word frequencies, and identified words' respective valences. This process attempts to uncover news outlets' positive or negative endorsements of certain moral dimensions concomitant with a particular ideology. In Experiment 1, the researchers gathered political articles from four sources. We were unable to reveal significant differences in moral or political endorsements, but we solidified the method to be employed in further research. In Experiment 2, the researchers expanded their number of sources to 10 and analyzed articles that pertain to two specific topics: the 2018 confirmation hearings of U.S. Supreme Court Justice Brett Kavanaugh and the partial U.S. Government Shutdown of 2018-2019. Once again, no significant differences in moral or political endorsements were found.

@@

Borrowing from Osgood's work in the 1950s, @Bradley1999 recognized valence as one of three related dimensions comprising emotion when developing their Affective Norms for English Words (ANEW). As mentioned before, "valence," the first dimension, denotes the pleasantness of a given word. "Arousal," the second dimension, describes the stimulating nature of a word. Lastly, "dominance" or "control" describes the extent to which a word makes one feel in or out of control [@Bradley1999]. The researchers developed ANEW by presenting participants with a list of 100-150 words and asking for them to rate the word on all three dimensions using the Self-Assessment Manikin (SAM), which allows ratings along either a nine-point scale when using traditional paper instruments or a twenty-point scale when using a computerized version.

@@

##import the overalldata 
overalldata = read.csv("../exp1_data/overalldata.csv")

@@

```{r data-cleanup, eval=FALSE, include=FALSE}
##pull in the data
master = read.csv("../exp1_data/overalldata.csv", stringsAsFactors = F)

@@

final = read.csv("../exp1_data/finaldata.csv", stringsAsFactors = F)
final = na.omit(final) #one weird NA line

@@

```{r stem_warriner, include = FALSE}
##open the Warriner data for later
warriner = read.csv("../exp1_data/BRM-emot-submit.csv", stringsAsFactors = F) #want v.mean.sum
warriner$V.Mean.Sum = scale(warriner$V.Mean.Sum)

@@

```{r create_datasets_harm, echo = F, include=FALSE}
#read in the harm total words from the mtmm project
harm_all = read.csv("../exp1_data/harm_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_fair, include=FALSE}
fair_all = read.csv("../exp1_data/fair_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_ingroup, include=FALSE}
ingroup_all = read.csv("../exp1_data/ingroup_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_authority, include=FALSE}
authority_all = read.csv("../exp1_data/authority_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_purity, include=FALSE}
purity_all = read.csv("../exp1_data/purity_words_total.csv", stringsAsFactors = F)

@@

As seen in Table \@ref(tab:exp1-source-descriptives), the sources are similar in some aspects yet different in others. Valence appears to be slightly positive across all sources. The large standard deviations seem to indicate little to no presence of a difference in valence across sources. 
*The New York Times* published the greatest number of articles as well as total words. *Breitbart* featured the lowest number of articles, and *NPR* the lowest number of total words from all articles. Per individual article, however, *Breitbart* appears to feature the highest average number of words as well as unique words. Once again the standard deviations call into question any apparent differences between sources. Finally, *Fox News* articles had the lowest reading grade level on average, while *The New York Times* had the highest. This result might be attributable to the greater number of tokens in the average *New York Times* article compared to *Fox News*. The standard deviations for readability suggest the presence of a diverse array of articles for each source, ranging from low to high reading level. Large standard deviations suggest the sources feature a lot of overlap between them in their representation of scores.

@@

```{r data_clean_kav, eval = T, echo = F}
##combine all kav data into one dataset
#read it in
NYtimes_kav = read.csv("../exp2_data/NYtimes_kav.csv", stringsAsFactors = F, fileEncoding="UTF-8") #ok
NPR_kav = read.csv("../exp2_data/NPR_kav.csv", stringsAsFactors = F, fileEncoding="UTF-8") #ok
SLATE_kav = read.csv("../exp2_data/SLATE_kav.csv", stringsAsFactors = F, fileEncoding="UTF-8") #ok
HUFFPO_kav = read.csv("../exp2_data/HUFFPO_kav.csv", stringsAsFactors = F, fileEncoding="UTF-8") #ok
POLITICO_kav = read.csv("../exp2_data/POLITICO_kav.csv", stringsAsFactors = F, fileEncoding="UTF-8") #ok 
FOX_kav = read.csv("../exp2_data/FOX_kav.csv", stringsAsFactors = F) #ok 
BREITBART_kav = read.csv("../exp2_data/BREITBART_kav.csv", stringsAsFactors = F) #ok
RUSH_kav = read.csv("../exp2_data/RUSH_kav.csv", stringsAsFactors = F, fileEncoding="UTF-8") #ok 
BLAZE_kav = read.csv("../exp2_data/BLAZE_kav.csv", stringsAsFactors = F, fileEncoding="UTF-8") #ok 
HANNITY_kav = read.csv("../exp2_data/HANNITY_kav.csv", stringsAsFactors = F, fileEncoding="UTF-8") #ok

@@

```{r data_clean_gs, eval = T, echo = F}
##combine all gs data into one dataset
#read it in
NYtimes_gs = read.csv("../exp2_data/NYtimes_gs.csv", stringsAsFactors = F, fileEncoding = "UTF-8") #ok
NPR_gs = read.csv("../exp2_data/NPR_gs.csv", stringsAsFactors = F) #ok
SLATE_gs = read.csv("../exp2_data/SLATE_gs.csv", stringsAsFactors = F) #ok 
HUFFPO_gs = read.csv("../exp2_data/HUFFPO_gs.csv", stringsAsFactors = F) #ok 
POLITICO_gs = read.csv("../exp2_data/POLITICO_gs.csv", stringsAsFactors = F, fileEncoding = "UTF-8") #not ok 
FOX_gs = read.csv("../exp2_data/FOX_gs.csv", stringsAsFactors = F) #ok 
BREITBART_gs = read.csv("../exp2_data/BREITBART_gs.csv", stringsAsFactors = F) #ok 
RUSH_gs = read.csv("../exp2_data/RUSH_gs.csv", stringsAsFactors = F, fileEncoding = "UTF-8") #not ok 
BLAZE_gs = read.csv("../exp2_data/BLAZE_gs.csv", stringsAsFactors = F) #ok 
HANNITY_gs = read.csv("../exp2_data/HANNITY_gs.csv", stringsAsFactors = F) #ok

@@

```{r create_datasets_harm_kav, eval = T, echo = F, include = F}
#read in the harm total words from the mtmm project
harm_all_kav = read.csv("../exp1_data/harm_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_fair_kav, include=FALSE}
fair_all_kav = read.csv("../exp1_data/fair_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_ingroup_kav, include=FALSE}
ingroup_all_kav = read.csv("../exp1_data/ingroup_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_authority_kav, include=FALSE}
authority_all_kav = read.csv("../exp1_data/authority_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_purity_kav, include=FALSE}
purity_all_kav = read.csv("../exp1_data/purity_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_harm_gs, eval = T, echo = F, include = F}
#read in the harm total words from the mtmm project
harm_all_gs = read.csv("../exp1_data/harm_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_fair_gs, echo = F, include=FALSE}
fair_all_gs = read.csv("../exp1_data/fair_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_ingroup_gs, echo = F, include=FALSE}
ingroup_all_gs = read.csv("../exp1_data/ingroup_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_authority_gs, echo = F, include=FALSE}
authority_all_gs = read.csv("../exp1_data/authority_words_total.csv", stringsAsFactors = F)

@@

```{r create_datasets_purity_gs, echo = F, include=FALSE}
purity_all_gs = read.csv("../exp1_data/purity_words_total.csv", stringsAsFactors = F)

@@

The results obtained in both experiments, while not confirmatory, provide several avenues for future investigation. Twice confronted with the failure to uncover differences in news media's political content, the researchers were unable to confirm the existence of an effect of political lean on American news writing under the auspices of Moral Foundations Theory. This does not, however, rule out the possibility that news media is biased. As @Mitchell2014 has demonstrated, self-reported political partisans are drawn to separate sources, but the present research was unable to offer any content-based explanations of this phenomenon. Even when attenuating the scraping criteria to collect text concerning (what the researchers believed to be) morally resonant topics, the method failed to detect differences in MFD word use by prominent news sources. Given part of the researchers' goal with this study was to answer a method question, they must contemplate whether or not the method or the construct of interest (political lean) are the primary contributor to null results. Either the method used to detect lean was insufficient for this purpose, or political lean simply does not exist, indicating the presence of a true negative in the results.

The researchers believe these results represent a false negative rather than a true negative. That is, we believe political lean exists, but that we have been unable to detect it at a content level. The researchers again turn to @Mitchell2014's work in this area: political partisans prefer certain news outlets over others. There must be something about the sources' content that facilitates partisans' repeated usage. Otherwise, they might not express such loyalty to any one source (or group of sources) over others. Given this aspect of the news economy, the researchers still believe certain news outlets in the U.S. contain a political slant. The researchers' focus, therefore, turns toward the limitations of the method, especially the chosen theoretical framework, under which the present research was conducted.

@@

There remains an aspect of language which this method still cannot address: context. Owing to the presence of millions of words in the analyses presented in the current study, the researchers would be unable to check for context and correct endorsement scores accordingly. @Graham2009 actually performed a process of human error-checking for context when formulating the MFD. While this might have helped that team fix mistakes made by the LIWC program (@Pennebaker2007), manual coding of this nature has the potential to introduce an unacceptable level of bias in the method, as perceptions of context would be dependent upon raters' interpretations and belief systems. Additionally, the numerous ways in which context affects meaning make it difficult to interpret MFD word occurrence outside of a narrow operational definition of "endorsement." In fact, political lean may be communicated through methods other than through the use of particular words.

One method through which news organizations might appeal to political partisans is through their content strategy rather than their use of MFD words. The results in the current study demonstrate a similar pattern of endorsement across both conservative and liberal sources. Specifically, the researchers found slightly positive endorsements for all five foundations no matter the political lean of the source. While this may simply be an indication of insufficient MFD word occurrence in news articles, it may also demonstrate truly equal endorsement despite political lean. While news organizations may endorse all moral foundations, they may also adopt strategies for doing so that appeal to the political partisans consuming their product. As an example, a liberal source might endorse *purity/sanctity* through an appeal to reduce waste or pollution (i.e. keeping the earth "clean"). A conservative source, on the other hand, might endorse *purity/sanctity* through an argument in favor of abstinence or religious commitment (i.e. keeping a "clean" soul, mind, or body). Both publications would register as positive endorsement of *purity/sanctity* despite appropriating the concept for distinct purposes. Through these divergent strategies, liberal and conservative news organizations might be able to appeal to their fans while exhibiting similar levels of moral endorsement. This could represent a further weakness of the Moral Foundations Dictionary as an instrument, as it is unable to identify these potential strategies for communicating certain points of view.

@@

@Belisle2018 conducted a text analysis (using a computer algorithm) of several speeches made by four recent U.S. presidents under Relational Frame Theory. These speeches, representing prominent examples of political discourse, were found to differ in their use of relational frames. For example, speeches by Donald Trump were found to contain more distinctive frames than other presidents, while Barack Obama utilized more coordinative frames than his counterparts. According to @Belisle2018, these observations imply President Obama, who employed more coordinative frames, utilized language emphasizing unity and similarity between people. Donald Trump, on the other hand emphasized differences and divisions between people through his use of distinctive frames [@Belisle2018]. At first glance, the differences appear to fall along party lines, though other factors may be involved as well. In any case, this research appears to indicate that RFT might be useful for analyzing political news content similar to the method employed in the current study.

@@

```{r appendix_mfd, results = 'asis', echo = FALSE, message = F}
orig_mfd = read.csv("../exp1_data/original_mfd.csv", stringsAsFactors = F, fileEncoding="UTF-8")
####build the table here####
tableprint = matrix(NA, nrow=length(orig_mfd), ncol=5)
colnames(tableprint) = c("Harm/Care", "Fairness/Reciprocity", "Ingroup/Loyalty", "Authority/Respect", "Purity/Sanctity")
